Exception Handling Bug Hazards in Android

### 2015-10-08 SPRINGER_JOURNAL/msr_journal.tex
including more quotes re-writting text

%////// PREVENT NULL POINTER /////
Developers were also asked about what they usually do to prevent NullPointerExceptions from happening when their app is in production. Most of them mentioned including null-checks in the application code (59\%), mostly when checking the return of a method and as method guard conditions. Many respondents (39\%) also answered that instead of just sprinkling the code with null-checks when a NullPointerException happens they investigate the cause of the null reference and work to fix it, instead of just adding a null check. Some respondents also mentioned using annotations such as @Nullable @NonNull, provided in Android since version XXX; such annotations guide the developer about what can and cannot be null during the application lifecycle. Although catching the NullPointerExceptions is considered a bad practice some respondents said that they do it to prevent crashes caused by NullPointerExceptions. Such behavior was mentioned by one of the respondents as the behavior adopted by non-experienced developers  to prevent such crashes -- which is undoubtedly a mistake. Some respondents also mentioned to replace the NullPointerException by IllegalArgumentException when a method should not receive a null as parameter. This is a good exception handling practice mentioned by Bloch [ref]. Only a few developers mentioned coding standards and static code analyzers to prevent it (2\%).

\greybox{``Determine why the object was null and attempt to fix this situation. In the case of external API calls which return null, then check for null (the quick-and-dirty way). For internal calls, use @NotNull and @Nullable annotations to provided more guidance on when an object \"may be\" and \"should never be\" null." [D42]}

\greybox{``Probably the most common exception, as such the reasons and fixes are super varying. But in general, null checks should be utilized and illegal state or argument exceptions should be used with an appropriate message. Which will communicate useful information without the confusion an NPE would normally cause."[D48]}

>>>>

%////// PREVENT NULL POINTER /////
Developers were also asked about what they usually do to prevent NullPointerExceptions from happening when their app is in production. Most of them mentioned including null-checks in the application code (59\%), mostly when checking the return of a method and as method guard conditions. Many respondents (39\%) also answered that instead of just sprinkling the code with null-checks when a NullPointerException happens they investigate the cause of the null reference and work to fix it, instead of just adding a null check. \emph{Determine why the object was null and attempt to fix this situation. In the case of external API calls which return null, then check for null (the quick-and-dirty way). For internal calls, use @NotNull and @Nullable annotations to provided more guidance on when an object \"may be\" and \"should never be\" null." [D42]}

Some respondents also mentioned using annotations such as @Nullable @NonNull, provided by Android Studio 0.5.5; such annotations guide the developer about what can and cannot be null during the application lifecycle. Although catching the NullPointerExceptions is considered a bad practice some respondents said that they do it to prevent crashes caused by NullPointerExceptions. Such behavior was mentioned by one of the respondents as the behavior adopted by non-experienced developers  to prevent such crashes -- which is undoubtedly a mistake. Some respondents also mentioned to replace the NullPointerException by IllegalArgumentException when a method should not receive a null as parameter. \emph{``Probably the most common exception, as such the reasons and fixes are super varying. But in general, null checks should be utilized and illegal state or argument exceptions should be used with an appropriate message. Which will communicate useful information without the confusion an NPE would normally cause."[D48]}.This is a good exception handling practice mentioned by Bloch [ref]. Only a few developers mentioned coding standards and static code analyzers to prevent it (2\%).
